---
title: "Unit 1 R Markdown"
author: "Bivin"
date: "2/20/2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# UNIT 1 R Review Part 1.R
## Manipulation and Import

### R Data Manipulation and Data Import
```{r}

#Vectors
age = c(22,21,24,19,20,23)
age
age[2]
yrs_math_ed = c(4,5,2,5,3,5)
yrs_math_ed

#data frames
df1 = data.frame(Age = age, Years = yrs_math_ed)
df1
df1[4,2] #Data in the 4th row and 2nd column
df1[1,] #all the data in the 1st row
df1$Math_Level #all the data in the fourth column

```

### Add a two more columns (variables).
```{r}
a = c("Mary","Martha","Kim","Kristen","Amy","Sam")
b = c("English","Math","Sociology","Math","Music","Dance")

#Make a data frame
df2 = data.frame(Name = a, Major = b)
df2
```

### Put it together 
```{r}
#Make another data frame
df3 = data.frame(Age = age, Years = yrs_math_ed,Name = a, Major = b)
df3
```

### Doing it a different way
```{r}
#Cbind combines columns
df4 = cbind(df1,df2)
df4
```

### Checking variable classes
```{r}
#str and class show what type of variable it is
str(df4)
class(df4$Name)
summary(df4)
```

### Define and try and add a new row (student)
```{r}
#define a new row / student
d = c(19,4,"John","Math")

#attempt to add a row
df5 = rbind(df4,d) #problem here is that Name and Major are factors
```


### Try and fix the problem.
```{r}
#try and fix the problem
dfCopy = df4
dfCopy$Name = as.character(df4$Name)
dfCopy$Major = as.character(df4$Major)
summary(dfCopy)
```

### Try and add the student again.
```{r}
#add the student to the end of the data frame.
df5 = rbind(dfCopy,d)
df5
```

### Look at the classes again.
```{r}
#recheck the class
str(df5)
summary(df5)
```

### Return/assign the right class to each variable.
```{r}
#fix the classes 
df5$Age = as.numeric(df5$Age)
df5$Years = as.numeric(df5$Years)
summary(df5)
df5$Name = as.factor(df5$Name)
df5$Major = as.factor(df5$Major)
summary(df5)
```

### Filtering
```{r}
#Filter the data frame
#All students with more than 4 years of Math
df5[df5$Years > 4,]
#All students with more than 4 years of Math and are 21 years of age or older
df5[(df5$Age >= 21 & df5$Years > 4), ]
#All students that are majoring in Math
df5[df5$Major == "Math",]
```

### Data Import from csv
```{r}
Example1 = read.csv("/Users/bivin/Desktop/OLD COMPUTER ARCHIVES/KadAfrica/MSDS/DDS/DDS2/Unit 1 Git-ing Sta-R-ted/Datasets/BusinessSales.csv",header = TRUE)
Example1$sales

```

### Pre loaded datasets / package "datasets" comes with the base R.

```{r}
# iris is Edgar Anderson's Iris Data
str(iris) # get type and other variable information
head(iris) #look at first 6 rows
tail(iris) #look at last 6 rows.
iris$Sepal.Width # look at all Sepal.Widths
```



### Data that comes with external packages
mpg is a dataset that comes in the **ggplot2** package, you may have to laod it with "data(mpg)"
mpg Fuel economy data from 1999 and 2008 for 38 popular models of car
```{r}
#install.packages("ggplot2") # I have already installed this package... you only need install the package once.
library(ggplot2) #However, you need to laod the library each time you start a new session.
?mpg #mpg info and note that it is a tibble. 
str(mpg) # get type and other variable information
head(mpg) #look at first 6 rows.
mpg$cty #look at all city mpg values.
```


# UNIT 1 R Review Part 2.R
## Basic "plot" functions: 

### Read in file.
```{r}
Example1 = read.csv("/Users/bivin/Desktop/OLD COMPUTER ARCHIVES/KadAfrica/MSDS/DDS/DDS2/Unit 1 Git-ing Sta-R-ted/Datasets/BusinessSales.csv",header = TRUE)
head(Example1)
```

### Scatterplots
```{r}
#plot(x,y,col,pch,type,ylab,xlab,main)
plot(Example1$ad_tv,Example1$sales, pch = 15,xlab = "TV Advertising $$$",ylab = "Sales $$$", main = "Sales v. Advertising")
abline(h = 55, col = "red",lwd = 5)
```

### plotting wth mpg
```{r}
plot(mpg$hwy,mpg$cty,pch = 15) #NO LABLES ... AHHH!!!
plot(mpg$hwy,mpg$cty,pch = 15, main = "City MPG v. Highway MPG", ylab = "City MPG", xlab = "Highways MPG")
```

### plot with Iris Data: Sepal Legth v. Petal Length
```{r}
plot(iris$Sepal.Length,iris$Petal.Length) #note domain and range of plot
```

### plot only viginica
```{r}
irisVir = iris[iris$Species == "virginica",]
plot(irisVir$Sepal.Length,irisVir$Petal.Length, col = "blue", ylim = c(0,7), xlim = c(4,8))
irisVers = iris[iris$Species == "versicolor",]
points(irisVers$Sepal.Length,irisVers$Petal.Length, col = "Red")
irisSet = iris[iris$Species == "setosa",]
points(irisSet$Sepal.Length,irisSet$Petal.Length, col = "green")
```

### histograms
```{r}
hist(mpg$cty,col = "blue")
```

### boxplots
```{r}
boxplot(mpg$cty~mpg$class, main = "Boxplot: City MPG v. Class")
```

### clearing the plot space
```{r}
dev.off()
```

### Dividing the plot space!  c(num rows, num columns)
```{r}
par(mfrow = c(1,2))

hist(mpg$cty,col = "blue", main = "Histogram of MPG", xlab = "MPG")
boxplot(mpg$cty~mpg$class, data= mpg, main = "Boxplot of MPG by Class", xlab = "Cylinder")
```

### You Try it! Make a histogram of the Iris Sepal Lengths
### Comment on the distribution (skewness?  number of modes? etc.)


```{r}
hist(iris$Petal.Length, col = "blue", main = "Histogram of Iris Sepal Lengths")
```

### barplot
```{r}
age = c(22,21,24,19,20,23)
yrs_math_ed = c(4,5,2,5,3,5)
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")

df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)

barplot(df3$Years, names.arg = df3$Name)
```

### We want a barplot of frequency of each class of car.
###note here how the data is in a different format for mpg.
###it is not one row per relevant observation with the value at the end
###we need to actually count the number of each class and then plot

### We could count ... that would take a while

### We could turn class into a factor and then let summary count for us.
```{r}
summary(mpg$class)

mpg$classFact = as.factor(mpg$class)
head(mpg)
summary(mpg$classFact)

barplot(summary(mpg$classFact))
```

# UNIT 1 R Review Part 3.R
## Statistics in R

### Sampling Distribution of the Individual
```{r}
#draw a sample from a standard normal distribution
#run many times varying sample size and look at histogram and mean
sample1 = rnorm(1000,0,1)
hist(sample1)
mean(sample1)
sd(sample1)
```

### Another way to code the same thing
```{r}
population = rnorm(10000000,0,1) #note the the number of draws here
hist(population)
sample1 = sample(population,100) #sample of size 100
hist(sample1)
mean(sample1)
sd(sample1)
```

### Sampling Distribution of the Sample Mean
now make a function in which you pass the sample size, the number of samples to take #and thus the number of sample means to return as well as the mean and sd.  

Define these outside the function so that they exist outside the function (global)

```{r}
xBarVec = c() #Global vector to hold the sample means
population = rnorm(10000000,0,1) #Simulating the population
```

### The Function
```{r}
#####################################################
# Funciton: xbarGenerator
# Argements: samplesize: the size of the sample that each sample mean is based on. 
#            number_of_samples: the number of samples and thus sample means we will generate
# Author: Bivin Sadler
#####################################################
xbarGenerator = function(sampleSize = 30,number_of_samples = 100)
{
  for(i in 1:number_of_samples)
  {
    theSample = sample(population,sampleSize)
    xbar = mean(theSample)
    xBarVec = c(xBarVec, xbar)
  }
  return(xBarVec)
}
```

### Before you Run This.... What do you think will be the "Sampleing Distribution of the Sample Mean"?
```{r}
xbars = xbarGenerator(30,1000)
length(xbars)
hist(xbars)
```

## You Try It
Now adapt the code above to generate a specified number of sample means of a specified size sample size (this has already been done) from a normally distributed population with a specified meanand standard deviation (currently it is forced to be from a standard normal distribuion N(0,1)).

Run it a few times and use it generate 10000 sample means, each from a sample of size 50, from a normal distribution with mean = 60 and sd = 10.  Provide a distribution (histogram) of the distrubtions as well. 

What is the shape of the distribuion?  What is the mean and sd?  What should they be given the Central Limit Theorem (CLT).  For a refresher of the CLT see the Stats Bridge Course!



### Possibe Solution
```{r}
xBarVec = c() #global vector to hold the sample means
#####################################################
# Funciton: xbarGenerator (Adpated)
# Argements: samplesize: the size of the sample that each sample mean is based on. 
#            number_of_samples: the number of samples and thus sample means we will generate
# Author: Bivin Sadler
#####################################################
xbarGenerator2 = function(sampleSize = 30,number_of_samples = 100, mean = 0, sd = 1)
{
  for(i in 1:number_of_samples)
  {
    theSample = rnorm(sampleSize,mean,sd)
    xbar = mean(theSample)
    xBarVec = c(xBarVec, xbar)
  }
  return(xBarVec)
}

xbars = xbarGenerator2(60,1000,50,10)
hist(xbars)
summary(xbars)
sd(xbars)
10/sqrt(60)
```
