---
title: "Unit 10 Overall"
author: "Bivin"
date: "4/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Unit 10 R Code

```{r}
library(tidyverse)
library(reshape2)

#Toy data set with miles per gallon versus weight
dfMPG = data.frame(MPG = c(42,40,38,28,32,30,18,20,22), Weight = c(1,1,1,2,2,2,3,3,3))

dfMPG %>% ggplot(aes(y = MPG, x = Weight)) + geom_point()

# set the yintercept and slope so we can change later to see the effects on the SSE
yint = 50.1
slope = -10

#Create the residual: observed - expected
residuals = dfMPG$MPG - (yint+dfMPG$Weight*slope)

#Look at sum of the residuals ( which is zero for the regression line)
sum(residuals)
#Look at the sum of the squared residuals which is the SSE or SSR
sum(residuals^2)

#Flexible Plotting
startx = 1
starty = yint + slope * startx
endx = 3
endy = yint + slope * endx

dfMPG %>% ggplot(aes(Weight,MPG)) + 
  geom_point() + 
  geom_segment(aes(x = startx, y = starty, xend = endx, yend = endy)) +
  xlim(0,4)
```


### Parameter Estimte Table and Distributions of Sample Intercept and Slope
```{r}
#Toy data set with miles per gallon versus weight
dfMPG = data.frame(MPG = c(42,40,38,28,32,30,18,20,22), Weight = c(1,1,1,2,2,2,3,3,3))

dfMPG %>% ggplot(aes(y = MPG, x = Weight)) + geom_point()

#Show lm() and estimation of toy example we have been using
fit = lm(MPG~Weight, data = dfMPG)
summary(fit)
#note that here beta_0_hat and beta_1_hat are beta_0 and beta_1
fit$coefficients
#beta_0_hat
fit$coefficients[1]
#beta_1_hat
fit$coefficients[2]

```

## Simulating Slope / By hand and with a loop
```{r}
### now lets simulate more samples from distributions described by beta_0 = 50 and beta_1 = -10 and sigma = 2
### When Weight = 1 (1000 lbs) Mu = 40
simWeight1 = rnorm(3,40,2)
### When Weight = 2 (2000 lbs) Mu = 30
simWeight2 = rnorm(3,30,2)
### When Weight = 3 (3000 lbs) Mu = 20
simWeight3 = rnorm(3,20,2)

### make the coresponding weights
weights = c(1,1,1,2,2,2,3,3,3)

### make a data frame so we can use lm() and ggplot()
dffullSample = data.frame( Weight = weights, MPG = c(simWeight1,simWeight2,simWeight3))

dffullSample

### plot the data
dffullSample %>% ggplot(aes(x = Weight, y = MPG)) + geom_point()

### plot the data and the line
dffullSample %>% ggplot(aes(x = Weight, y = MPG)) + geom_point() + geom_smooth(method = "lm")

### get the sample intercept and slope
fit = lm(MPG~Weight, data = dffullSample)
#summary(fit)
fit$coefficients[1]
fit$coefficients[2]
```


## Loop for simulating many sample and thus many sample intercepts
```{r}
#now let's put it in a loop and see if we can see the distribution of the intercept
# remember that mu = 50-10*Weight and sigma = 2

#Holds the generated sample intercepts
sampInterceptHolder = c()
numSampIntercepts = 10000
for (i in 1:numSampIntercepts)
{
  #now lets simulate more samples from distributions described by beta_0 = 50 and beta_1 = -10 and sigma = 2
  #When Weight = 1 (1000 lbs) Mu = 40
  simWeight1 = rnorm(3,40,2)
  #When Weight = 2 (2000 lbs) Mu = 30
  simWeight2 = rnorm(3,30,2)
  #When Weight = 3 (3000 lbs) Mu = 20
  simWeight3 = rnorm(3,20,2)
  
  # make the coresponding weights
  weights = c(1,1,1,2,2,2,3,3,3)
  
  #make a data frame so we can use lm() and ggplot()
  dffullSample = data.frame(Weight = weights, MPG = c(simWeight1,simWeight2,simWeight3))
  
  #get the sample intercept and slope
  fit = lm(MPG~Weight, data = dffullSample)
  sampInterceptHolder[i] = fit$coefficients[1]
}

#Used a simple histogram (rather than from ggplot) since not for presentation
hist(sampInterceptHolder, main = str_c("Distribution of Intercpets from ", numSampIntercepts, " Samples"), xlab = "Sample Intercpet")
summary(sampInterceptHolder)
```

## Now show calculation of t-statistic and pvalue
```{r}
fit = lm(MPG~Weight, data = dfMPG)
summary(fit)
confint(fit)

#Intercept
tstat = 50/1.6330 #beta_0_hat / SE(beta_0_hat)
pvalue = (1-pt(tstat,7)) * 2 # Mult by 2 since 2 sided test
tstat
pvalue

#Slope
tstat = -10/.7559 #beta_0_hat / SE(beta_0_hat)
pvalue = (pt(tstat,7)) * 2 # Mult by 2 since 2 sided test
tstat
pvalue

# Back to slides for hypothesis test
```


###Example: MPG v. Weight (MPG R Dataset)
```{r}
mtcars2 = mtcars %>% filter(wt < 4.5)
fit = lm(mpg~wt, mtcars2)
summary(fit)
confint(fit)
mercedes = data.frame(wt = 2.304)
predict(fit, newdata = mercedes, interval = "confidence")
```

###Transformations.

```{r}
#Recall

mtcars %>% 
  ggplot(aes(x = wt, y = mpg)) + geom_point() + ggtitle("mtcars: mpg v. weight") + geom_smooth(method = "lm") + xlim(0,6)

# degree 1 model
fit = lm(mpg~wt, data = mtcars)
summary(fit)

#degree 2 model
mtcars %>% ggplot(aes(x = wt, y = mpg)) + geom_point()
mtcars3 = mtcars %>% mutate(wt2 = wt^2)
fit = lm(mpg~wt+wt2, mtcars3)
summary(fit)

preds = predict(fit)
mtcars %>% ggplot(aes(x = wt, y = mpg)) + geom_point() +geom_line(data = mtcars, aes( x = wt, y = preds, col = "red"))

#deg 3 model 
mtcars %>% ggplot(aes(x = wt, y = mpg)) + geom_point()
mtcars3 = mtcars %>% mutate(wt2 = wt^2, wt3 = wt^3)
fit = lm(mpg~wt+wt2+wt3, mtcars3)
summary(fit)
preds = predict(fit)
mtcars %>% ggplot(aes(x = wt, y = mpg)) + geom_point() +geom_line(data = mtcars, aes( x = wt, y = preds, col = "red"))
```


### MLR plotly and 3D plotting
```{r}
fit = lm(mpg ~ wt + hp, data = mtcars)
summary(fit)

preds = predict(fit)
preds
mtcars$predMPG = preds

#Estimate the mpg for a car that weights 2000lbs and has 200 hp.  
df = data.frame(wt = 2, hp = 200)
predict(fit, newdata = df)

#plot the regression plane

#Graph Resolution (more important for more complex shapes)
graph_reso <- .5

#Setup Axis
axis_x <- seq(min(mtcars$hp), max(mtcars$hp), length = 100)
axis_y <- seq(min(mtcars$wt), max(mtcars$wt), length = 100)

#Sample points
surfaceMPG <- expand.grid(wt = axis_y,hp = axis_x,KEEP.OUT.ATTRS = F)
surfaceMPG$mpg <- predict.lm(fit, newdata = surfaceMPG)
surfaceMPG <- acast(surfaceMPG, wt ~ hp, value.var = "mpg") #y ~ x

surface_plot <- plot_ly(mtcars, 
                     x = ~hp, 
                     y = ~wt, 
                     z = ~mpg,
                     text = "mpg", 
                     type = "scatter3d",
                     mode = "markers")


surface_plot <- add_trace(p = surface_plot,
                       z = surfaceMPG,
                       x = axis_x,
                       y = axis_y,
                       type = "surface")

surface_plot

```



###Plot of assumptions being met for MLR in 3D. and Curse of Dimensionality
```{r}
#Explanatory Variables
axis_x <- seq(1,10, by = 1)
axis_y <- seq(1,10, by = 1)
GridOfXs <- expand.grid(x = axis_x,y = axis_y,KEEP.OUT.ATTRS = F)

df = data.frame()

#Create coordinates for a random sample of points on the grid
set.seed(22) #for repeatability
for (i in 1 : 6) #vary the last argument  and make last argument in sample call below 1 for independent observations.
{
  x = rep(sample(1:10,1),100) #how many we are pulling for each xy pair.
  y = rep(sample(1:10,1),100) #how many we are pulling for each xy pair.
  
  z = 4 + 3*x - 7*y + rnorm(100,0,7) #make sure the you generate the same number of errors as the last argument in sample call. 
  
  df = rbind(df, data.frame(x = x, y = y, z = z))
}

surface_plot <- plot_ly(df, 
                        x = ~x, 
                        y = ~y, 
                        z = ~z,
                        text = "z", 
                        type = "scatter3d",
                        mode = "markers")

surface_plot

#Get coordinates for all points
AllZs = 4 + 3*GridOfXs[,1] - 7*GridOfXs[,2]
Coords = cbind(GridOfXs,AllZs)

#Put in format for plotting with plot_ly
CoordsMat <- acast(Coords, y ~ x, value.var = "AllZs") #y ~ x

surface_plot <- add_trace(p = surface_plot,
                          z = CoordsMat,
                          x = axis_x,
                          y = axis_y,
                          type = "surface")

surface_plot


# Histogram and Scatter plot of Residuals

fit = lm(z~x + y, data = df)
hist(fit$residuals, col = "blue", main = "Histogram of Residuals")
plot(fit$fitted.values,fit$residuals, main = "Plot of Residuals v. Fitted Values")

library(GGally)
ggpairs(df)
```


# Assumptions for MPG = Weigtht + HP MLR
```{r}
library(reshape2)

#Graph Resolution (more important for more complex shapes)
graph_reso <- .5

#Setup Axis
axis_x <- seq(min(mtcars$hp), max(mtcars$hp), length = 100)
axis_y <- seq(min(mtcars$wt), max(mtcars$wt), length = 100)

#Sample points
surfaceMPG <- expand.grid(wt = axis_y,hp = axis_x,KEEP.OUT.ATTRS = F)
surfaceMPG$mpg <- predict.lm(fit, newdata = surfaceMPG)
surfaceMPG <- acast(surfaceMPG, wt ~ hp, value.var = "mpg") #y ~ x

surface_plot <- plot_ly(mtcars, 
                        x = ~hp, 
                        y = ~wt, 
                        z = ~mpg,
                        text = "mpg", 
                        type = "scatter3d",
                        mode = "markers")


surface_plot <- add_trace(p = surface_plot,
                          z = surfaceMPG,
                          x = axis_x,
                          y = axis_y,
                          type = "surface")

surface_plot

fit = lm(mpg~wt+hp, data = mtcars)
hist(fit$residuals, col = "blue", main = "Histogram of Residuals")
plot(fit$fitted.values,fit$residuals, main = "Plot of Residuals v. Fitted Values")

library(GGally)
mtcars %>% select(mpg,wt,hp) %>% ggpairs()
```



## Cross Validation
```{r}
NBA = read.csv(file.choose(),header = TRUE)
NBATrain = read.csv(file.choose(),header = TRUE)
NBATest = read.csv(file.choose(),header = TRUE)
Model1_fit = lm(APPG ~ Height+Weight+FG+FT, data = NBATrain)
summary(Model1_fit)
Model1_Preds = predict(Model1_fit, newdata = NBATest)
as.data.frame(Model1_Preds)

# New Cross Validation 
set.seed(4)
TrainObs = sample(seq(1,dim(NBA)[1]),round(.75*dim(NBA)[1]),replace = FALSE)
NBATrain = NBA[TrainObs,]
NBATrain
NBATest = NBA[-TrainObs,]
NBATest
Model1_fit = lm(APPG ~ Height+Weight+FG+FT, data = NBATrain)
summary(Model1_fit)
Model1_Preds = predict(Model1_fit, newdata = NBATest)
as.data.frame(Model1_Preds)

# Model 1
MSPE = data.frame(Observed = NBATest$APPG, Predicted = Model1_Preds)
MSPE$Resisdual = MSPE$Observed - MSPE$Predicted
MSPE$SquaredResidual = MSPE$Resisdual^2
MSPE
mean(MSPE$SquaredResidual)

#Alternative MSPE Calculation
MSPE = mean((NBATest$APPG - Model1_Preds)^2)
MSPE

#Model 2
Model2_fit = lm(APPG ~ Height + FG, data = NBATrain)
summary(Model2_fit)
Model2_Preds = predict(Model2_fit,newdata = NBATest)
as.data.frame(Model2_Preds)
MSPE = mean((NBATest$APPG - Model2_Preds)^2)
MSPE
```



## More Stable Measure ... Average of many MSPEs
```{r}
numMSPEs = 1000
MSPEHolderModel1 = numeric(numMSPEs)
MSPEHolderModel2 = numeric(numMSPEs)

for (i in 1:numMSPEs)
{
  TrainObs = sample(seq(1,dim(NBA)[1]),round(.75*dim(NBA)[1]),replace = FALSE)
  NBATrain = NBA[TrainObs,]
  NBATrain
  NBATest = NBA[-TrainObs,]
  NBATest
  Model1_fit = lm(APPG ~ Height+Weight+FG+FT, data = NBATrain)
  Model1_Preds = predict(Model1_fit, newdata = NBATest)
  
  #MSPE Model 1
  MSPE = mean((NBATest$APPG - Model1_Preds)^2)
  MSPE
  MSPEHolderModel1[i] = MSPE
  
  #Model 2
  Model2_fit = lm(APPG ~ Height + FG, data = NBATrain)
  Model2_Preds = predict(Model2_fit,newdata = NBATest)
  MSPE = mean((NBATest$APPG - Model2_Preds)^2)
  MSPE
  MSPEHolderModel2[i] = MSPE
  
}

mean(MSPEHolderModel1)
mean(MSPEHolderModel2)
```
