---
title: "Unit 2 Visualization"
author: "Bivin"
date: "4/28/2019"
output: html_document
---

## UNIT 2 R Quick Start / Scatterplot, histogram and bar plot.R

### Remember Previous Plots from Unit 1
### Let's redo them using ggplot2!

Scatterplot
```{r}
#install.packages("ggplot2")
#library(ggplot2)

#Iris Scatterplot
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Petal.Length)) + geom_point()

#With pipe!  %>% = "and then"
iris %>% ggplot(aes(x = Sepal.Length,y = Petal.Length)) + geom_point() #Labels but not TITLE! AHHH!!!!

iris %>% ggplot(aes(x = Sepal.Length,y = Petal.Length)) + geom_point() + ggtitle("City MPG v. Highway MPG")

# You Try It! (mpg)
# plot (scatter plot) city mile per gallon (y) versus highway miles per gallon (x).

mpg %>% ggplot(aes(x = hwy, y = cty)) + geom_point() + ggtitle("City MPG v. Highway MPG")
```


Histogram
```{r}
#Iris histogram
iris %>% ggplot(aes(x = Petal.Length)) + geom_histogram() + ggtitle("Histogram of Petal Length")


# You Try It!  investigate the distribuion of the city miles per gallon variable for compact cars.  

mpg[mpg$class == "compact",] %>% ggplot(aes(x = cty)) + geom_histogram()

#You Try It!  No compare the distribution of cty for compact cars to that of SUVs.  

mpg[mpg$class == "suv",] %>% ggplot(aes(x = cty)) + geom_histogram()

# You Try It Bonus!  In order to compare the distributions, we should put them on the same scale.
# Research how to change the scale!

mpg[mpg$class == "compact",] %>% ggplot(aes(x = cty)) + geom_histogram() + coord_cartesian(xlim = c(0, 50)) 
```

Boxplot
```{r}
#Iris Boxplot
iris %>% ggplot(aes(x = Species, y = Petal.Length)) + geom_boxplot()


# You Try It!

#Find a box plot from the mpg dataset of the city mpg versus the class of car.
mpg %>% ggplot(aes(x = class, y = cty)) + geom_boxplot() + ggtitle("Boxplot of City MPG v. Class")
```



### Bar plot with mpg and class
```{r}
#With base barplot
summary(mpg$class)

mpg$classFact = as.factor(mpg$class)
head(mpg)
summary(mpg$classFact)

barplot(summary(mpg$classFact))

# With ggplot /  geom_bar()
mpg %>% ggplot(aes(x = class)) + geom_bar() + ggtitle("Distribution of Class")
```

## Bar plot with toy Student and Math Experience data
```{r}
#With base barplot
age = c(22,21,24,19,20,23)
yrs_math_ed = c(4,5,2,5,3,5)
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")

df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)

barplot(df3$Years, names.arg = df3$Name)


# With ggplot /  geom_bar()   Where does Years go? We will come back to this!
df3 %>% ggplot(aes(x = Name)) + geom_bar() + ggtitle("Distribution of Class")

df3 %>% ggplot(aes(x = Name, y = Years)) + geom_bar(stat = "identity") + ggtitle("Distribution of Class")

df3 %>% ggplot(aes(x = Name, y = Years)) + geom_col() + ggtitle("Distribution of Class")



# You Try It!
# Iris Data:  Create a Barplot of how many of each Species are in the study.

iris %>% ggplot() + geom_bar(aes(x = Species))

# You Try it Bonus: Create a dataframe of how you rank (scale of 1(least) to 10 (best)) traveling by car, train, plane, boat (ship), and bike.

travelDF = data.frame(type = c("car", "train", "plane", "boat", "bike"), rank = c(7,9,8,5,6))

travelDF %>% ggplot(aes(x = type, y = rank)) + geom_col() + ggtitle("Tavel Ranking")
```



## AES
### AES: X and Y
```{r}
mpg %>% 
ggplot(mapping = aes(x = hwy, y = cty)) +
geom_point()

mpg %>% 
ggplot(mapping = aes(x = hwy)) + 
geom_histogram()

mpg %>% 
ggplot(mapping = aes(x = class)) + 
geom_bar()

mpg %>% + ggplot(mapping = aes(x = class, y = ..prop.., group = 1)) + geom_bar()
```

### AES: color
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty,color = mpg$drv)) + geom_point()

# color = "blue" is not a mapping of a variable
mpg %>% ggplot(aes(x = hwy, y = cty, color = "blue")) + 
geom_point()

#note that color is not in an ”aes”
# function we are setting them manually.
mpg %>% ggplot(aes(x = hwy, y = cty)) + 
geom_point(color = "blue")

# note now we are ”mapping” fill to a variable (drv).
mpg[mpg$class == "compact",] %>% ggplot(aes(x = cty, fill = drv)) + 
geom_histogram(color = "black") 

#note that fill and color are not in an ”aes” function we are setting them manually.
# We are not “mapping” variables to colors… rather just setting the color.  
mpg[mpg$class == "compact",] %>% ggplot(aes(x = cty)) + 
geom_histogram(fill = "blue", color = "black") 

# note now we are ”mapping” fill to a variable (drv).
mpg %>% ggplot(aes(x = class, y = cty,fill = drv)) + geom_boxplot(color = "blue")

#note that fill and color are not in an ”aes” function we are setting them manually.
# We are not “mapping” variables to colors… rather just setting the color.  
mpg %>% ggplot(aes(x = class, y = cty)) + geom_boxplot(color = "blue", fill = "black") 

```


### AES: size
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty, size = mpg$displ)) + geom_point()

```


### AES: size
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty, size = mpg$displ)) + geom_point()

```


### AES: size
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty, size = mpg$displ)) + geom_point()

```



### AES: alpha
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty, alpha = mpg$displ)) + geom_point()

```

### AES: shape
```{r}
mpg %>% ggplot(aes(x = hwy, y = cty, shape = mpg$drv)) + geom_point()

```


## GEOMS

### geom_smooth
```{r}
ggplot(data = mpg) + 
geom_point(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

```

### geom_smooth you try it
```{r}
ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty))

ggplot(data = mpg) + 
geom_smooth(mapping = aes(x = hwy, y = cty))

ggplot(data = mpg) + 
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = drv))



#Global aes() ... one aes that applies to all layers
ggplot(data = mpg, mapping = aes(x = hwy, y = cty, linetype = drv, color = drv)) + 
geom_point() +
geom_smooth()

#Individual aes() …. an aesthetic for each layer
ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty, color = class)) +
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = drv, color = drv))

#Individual aes() …. an aesthetic for each layer
ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty, color = class)) +
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = class, color = class)) + facet_wrap(~class)


```

### labels
```{r}
ggplot(data = mpg, aes(x = class, y = cty)) + 
geom_boxplot() + 
ggtitle("Boxplot of City MPG v. Class") + 
xlab("Class") + ylab("City MPG")
```


### facets
facet wrap
```{r}
ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty, color = class)) +
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = drv)) +
facet_wrap(~class)

ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty, color = drv)) +
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = drv)) +
facet_wrap(~class)


```

facet grid
```{r}
ggplot(data = mpg) + 
geom_point(mapping = aes(x = hwy, y = cty, color = drv)) +
geom_smooth(mapping = aes(x = hwy, y = cty, linetype = drv)) +
facet_grid(drv~class)

ggplot(data = mpg, aes(fill = drv)) + 
geom_histogram(mapping = aes(x = cty)) +
facet_grid(drv~class)

```

## Transformations
```{r}
mpg %>% ggplot(aes(x = class)) + geom_bar(stat = "count")

mpg %>% ggplot(aes(x = class, y = cty)) + geom_bar()

#stat = identity adds the values of the y value per class (x level)
mpg %>% ggplot(aes(x = class, y = cty)) + geom_bar(stat = "identity")

```

## Positions
### stack
```{r}
mpg %>% ggplot(mapping = aes(x = class, fill = drv)) + geom_bar(position = "stack")

mpg %>% ggplot(mapping = aes(x = class, fill = drv)) + geom_bar()

```

### dodge
```{r}
mpg %>% ggplot(mapping = aes(x = class, fill = drv)) + geom_bar(position = "dodge")

```

### fill
```{r}
mpg %>% ggplot(mapping = aes(x = class, fill = drv)) + geom_bar(position = "fill")

```

### jitter
```{r}
mpg %>% 
+ ggplot(mapping = aes(x = hwy, y = cty)) + 
+ geom_point()

mpg %>% 
+ ggplot(mapping = aes(x = hwy, y = cty)) + 
+ geom_point(position = "jitter")

```

## Other Coordinate Systems
### coord_flip
```{r}
mpg %>% + ggplot(mapping = aes(x = class, fill = class)) + geom_bar() + coord_flip()
```

### coord_polar
```{r}
mpg %>% + ggplot(mapping = aes(x = class, fill = class)) + geom_bar() + coord_polar()
```

### coord_quickmap and map_data
```{r}
install.packages("maps")
library()

usa = map_data("usa")

p <- ggplot() + 
  geom_polygon(data = usa, aes(x = long, y = lat, group = group), fill = "blue", color = "black") + 
  coord_quickmap()

#Dallas Coords
Dallas <- tibble(
  long = c(-96.7970),
  lat = c(32.7767),
  names = c("Dallas")) 

p + geom_point(data = Dallas, aes(x = long, y = lat), shape = 21, color = "black", fill = "yellow", size = 5) +
  geom_text(data = Dallas, aes(x = long, y = lat, label = names), hjust = 0, nudge_x = 1, color = "white")


#Italy Map
italy = map_data("italy")
ggplot(italy, aes(long,lat, group = group)) + 
geom_polygon(fill = "white", color = "black") +
coord_quickmap()
```

## themes
```{r}
install.package("ggthemes")
library(ggthemes)

#excel
ggplot(mapping = aes(x = hwy, y = cty)) + 
geom_point(position = "jitter") + 
ggtitle("City MPG v. Highway MPG") + 
theme_excel()

mpg %>% 
+ ggplot(mapping = aes(x = class, fill = class)) +  geom_bar() + theme_excel()

#Economist
ggplot(mapping = aes(x = hwy, y = cty)) + 
geom_point(position = "jitter") + 
ggtitle("City MPG v. Highway MPG") + 
theme_economist()

mpg %>% 
+ ggplot(mapping = aes(x = class)) +  geom_bar() + theme_economist()

#538
ggplot(mapping = aes(x = hwy, y = cty)) + 
geom_point(position = "jitter") + 
ggtitle("City MPG v. Highway MPG") + 
theme_fivethirtyeight()

mpg %>% +
ggplot(mapping = aes(x = class, fill = class)) +  
geom_bar() + ggtitle("Vehicle Class") + 
theme_fivethirtyeight()

```

## Plotly!!!
```{r}

#MPG
library(ggthemes)
library(plotly)

p = mpg %>% 
ggplot(mapping = aes(x = hwy, y = cty)) + 
geom_point(position = "jitter") + ggtitle("City MPG v. Highway MPG") + theme_excel()
 
ggplotly(p)


p = mpg[mpg$class == "compact",] %>% 
ggplot(aes(x = cty, fill = drv)) + 
geom_histogram(color = "blue") 

ggplotly(p)
```

Student Data
```{r}
#generating the data
age = c(22,21,24,19,20,23)
yrs_math_ed = c(4,5,2,5,3,5)
names = c("Mary","Martha","Kim","Kristen","Amy","Sam")
subject = c("English","Math","Sociology","Math","Music","Dance")
df3 = data.frame(Age = age, Years = yrs_math_ed, Name = names, Subject = subject)

#plotting code
p = df3 %>% 
ggplot(aes(x = Name, y = Years)) + 
geom_bar(stat = "identity") + 
ggtitle("Distribution of Class")

#enabling interactive piece
ggplotly(p)
```

## 3D Plots With Plotly
```{r}
#mpg
p <- plot_ly(mpg, x = ~cty, y = ~hwy, z = ~displ, color = ~drv) %>%
   add_markers() %>%
   layout(scene = list(xaxis = list(title = 'City MPG'),
                      yaxis = list(title = 'Displacement'),
                      zaxis = list(title = 'Highway MPG')))
p

# You Try... Create a 3D plot with plotly with x = sepal width, y = sepal.length, z = petal.width




p <- plot_ly(iris, x = ~Sepal.Width, y = ~Sepal.Length, z = ~Petal.Width, color = ~Species) %>%
   add_markers() %>%
   layout(scene = list(xaxis = list(title = 'Sepal Width'),
                      yaxis = list(title = 'Sepal.Length'),
                      zaxis = list(title = 'Petal.Width')))
p
```



