---
title: "Unit 5 Overall"
author: "Bivin"
date: "5/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Merging and Joins
```{r}
df1 = data.frame(Student_ID = c("1234", "2345", "8910", "9101", "3456", "5678","8888"), Course = c("Time Series", "NLP", "Stats1", "DDS", "DDS", "ML2","Data Mining"))
df2 = data.frame(Student_ID = c("1234", "2345", "8910", "9101", "3456", "5678","99999", "11111"), Gender = c("M", "F", "M", "F", "F", "F", "M", "M"), State = c("TX", "TX", "CA", "ID", "NY", "FL","NM", "AZ") )

#inner join
merge(df1,df2, by = "Student_ID")
#or
inner_join(df1,df2,by = "Student_ID")
#or
df1 %>% inner_join(df2,by = "Student_ID")


#outer join
merge(df1,df2, by = "Student_ID",all = TRUE)
#or
full_join(df1,df2,by = "Student_ID")
#or
df1 %>% full_join(df2,by = "Student_ID")


#left join
merge(df1,df2, by = "Student_ID",all.x = TRUE)
left_join(df1,df2, by = "Student_ID")


#right join 
merge(df1,df2, by = "Student_ID",all.y = TRUE)
right_join(df1,df2, by = "Student_ID")


#Different Column Names: by.x and by.y
df3 = data.frame(Student_ID_Number = c("1234", "2345", "8910", "9101", "3456", "5678", "8888"), Course = c("Time Series", "NLP", "Stats1", "DDS", "DDS", "ML2", "Data Mining"))
df4 = data.frame(Student_ID = c("1234", "2345", "8910", "9101", "3456", "5678","99999", "11111"), Gender = c("M", "F", "M", "F", "F", "F", "M", "M"), State = c("TX", "TX", "CA", "ID", "NY", "FL","NM", "AZ") )

merge(df3,df4, by.x = "Student_ID_Number", by.y = "Student_ID")
```





#Headmap from AcuSpike
```{r}
#Data clean and heat map for Acuspikes 
# Data has been read into a Dataframe called "Acu"

library(ggplot2)
library(maps)
library(dplyr)
library(mapproj)

Acu = read.csv(file.choose(),header = TRUE) # read in company data
lookup = data.frame(abb = state.abb, State = state.name) #makes a data frame with State name and abbreviation. 
colnames(Acu)[2] = "abb" # Change Column Name
Acu2 = merge(Acu,lookup,"abb") # make one dataset with state names and abb
AcuMapData = count(Acu2,State) #count up the occurance of each state. 
#AcuMapData = AcuMapData[-c(5,9,43),] #Shows contrast between other states better
colnames(AcuMapData)[2] = "AcuSpikes" #change "n" to "Acuspikes"
AcuMapData$region <- tolower(AcuMapData$State)
AcuMapData2 = AcuMapData[-1]
states <- map_data("state")
map.df <- merge(states,AcuMapData2, by="region", all.x=T)
map.df <- map.df[order(map.df$order),]
ggplot(map.df, aes(x=long,y=lat,group=group))+
  geom_polygon(aes(fill=AcuSpikes))+
  geom_path()+ 
  scale_fill_gradientn(colours=rev(heat.colors(10)),na.value="grey90")+ggtitle("Acuspike Systems by State")+
coord_map()
```


# WDI and WHO: GDP Per Capita Versus Life Expectancy
## Indicator API Queries:  
https://datahelpdesk.worldbank.org/knowledgebase/articles/898599-indicator-api-queries  
See the csv of Indicators as well.  
```{r}
install.packages("countrycode")
library(countrycode)
countrycode("Albania", 'country.name', 'iso2c') 
countrycode("Germany", 'country.name', 'iso2c') 
countrycode("United States", 'country.name', 'iso2c') 
countrycode("Canada", 'country.name', 'iso2c') 


library(WDI)
#GDP ("NY.GDP"), Per capita ("PCAP"), in US Currency (KD)
GDPPC = WDI(indicator="NY.GDP.PCAP.KD", start=2015, end=2015)
head(GDPPC)

library(WHO)
codes <- get_codes()
str(codes)
dim(codes)
names(codes)
grep("life expect",codes$display)
#str_which(codes$display,"life expect")
# another way instead of grep
#which(str_detect(codes$display,"life expect"))
codes$display[58]
codes$display[2187]
codes$label[58]
codes[58,]
#58th codes has label "WHOSIS_000002"
LE = get_data("WHOSIS_000002")
LE$iso2c = countrycode(LE$country, 'country.name', 'iso2c') 
LE
LE2 = LE %>% filter(sex == "Both sexes" & year == 2015)
LE2
GDPPC2 = GDPPC[48:264,]  #These are the rows of the indiviudal countries rather than regions
GDPPC3 = GDPPC2[!is.na(GDPPC2$NY.GDP.PCAP.KD),]
head(GDPPC3)
dim(GDPPC2)
dim(GDPPC3)
# Try and get a GDP for every country in the LE2 set. 
GDPPC_LE = left_join(LE2,GDPPC3,"iso2c") %>% select(country.x,NY.GDP.PCAP.KD,value,year.x)
head(GDPPC_LE)
#plot the data
library(plotly)
names(GDPPC_LE)[1] = "Country"
ThePlot = GDPPC_LE %>% 
  ggplot(aes(NY.GDP.PCAP.KD,value,color = Country)) + geom_point() + xlab("GDP Per Capita") +   ylab("Life Expectancy") + ggtitle("Life Expectancy v. GDP Per Capita for 2015")
ggplotly(ThePlot)
```


#String Basics
```{r}
string1 = "Don't let what you can't do interfere with what you can do."
string2 = "\"Don't let what you can't do interfere with what you can do.\" - John Wooden"
writeLines(string2)


string3 = c("IF","YOU","GET","GIVE,", "IF","YOU","LEARN","TEACH", "-MAYA", "ANGELOU")
string3

#strinr functions
str_length(string3)

#Collapse Strings... put them together
str_c(string3)
#Collapse Strings... put them together
str_c(string3, collapse = "")
#Collapse Strings... put them together
str_c(string3, collapse = ",")
#Collapse Strings... put them together
str_c(string3, collapse = " ")

string4 = str_c(string3, collapse = " ")
string4
str_sub(string4,1,1)
str_sub(string4,1,7)
str_sub(string4,-3,-1)
str_sub(string4,-14,-1)
str_sub(string4,3,7)
```


# Regular Expressions
## Basics
```{r}
string5 = " \"I've missed more than 9000 shots in my career. I've lost almost 300 games. 26 times, I've been trusted to take the game winning shot and missed. I've failed over and over and over again in my life. And that is why I succeed.\" - Michael Jordan"

str_view(string5, "shot") 
str_view_all(string5,"shot")
str_view_all(string5,"shot\\.")
str_view_all(string5,"\\bshot\\b") #Important
str_view_all(string5,"I")
str_view_all(string5," I ")
str_view_all(string5,"\\. I") #Find "I" at the beginning of a sentence. 
str_view_all(string5,"(\\. I| \"I)")
#How can we count how many times it comes up in the string?  The answer is coming.


str_view(string2, "do.") # . is the wildcard character 
str_view(string2, "do\\.") # need \\. because \ is an escape in both strings and regular expressions
```

##Anchors
```{r}
string6 = " \"Sometimes when you innovate, you make mistakes. It is best to admit them quickly, and get on with improving your other innovations.\" - Steve Jobs"
str_view(string1, "^Don't")
str_view(string1, "^interfere")
str_view(string3, "^T")
str_view(string6, "^\"Some") #Why Nothing ... we will come back to this.
str_view(string6, "Jobs$")

#Find digits and whitespace
str_view(string5,"\\d")
str_view_all(string5,"\\d")
str_view_all(string6,"\\s") #Ahh... there was white space
str_view(string6, "^ \"Some") 
```

## Character Classes and Alternatives
```{r}
#Say we have a list of quotes and we want to find those from Steve Jobs
#But we don't know if it is under Steve, Jobs or Steve Jobs  
#So we look for Steve or Jobs

df = data.frame(quotes = c(string2,string4,string5,string6))
df$quotes
str_view(df$quotes,"(Steve|Jobs)")
str(df) # quotes are factors by default on this machine.

df = data.frame(quotes = c(string2,string4,string5,string6), stringsAsFactors = FALSE)
df$quotes
str_view(df$quotes,"(Steve|Jobs)")
str_view_all(df$quotes,"[abc]")
str_view_all(df$quotes,"[^abc]")
```

## Repitition
```{r}
# ?: 0 or 1
# +: 1 or more
# *: 0 or more

str_view(df$quotes, "0+") #Greedy
str_view(df$quotes, "[sn]+") #Greedy
str_view_all(df$quotes, "[sn]+")

#More Repitions
# {n}: exactly n
# {n,} n or more
# {,m} at most m

str_view_all(df$quotes, "[sn]{2}") #Greedy
```



## Detect and Count ... stringr and base
```{r}
#Recall
str_view_all(string5,"shot")

str_detect(string5,"shot")
str_detect(string5,"rainbow")


str_count(string5,"shot")
str_count(string5,"rainbow")


#from earlier
str_view_all(string5,"(\\. I| \"I)")
str_detect(string5,"(\\. I| \"I)")
str_count(string5,"(\\. I| \"I)")

#grepl
grepl("and",string5) #similar to str_detect


str_detect(df$quotes,"shot")
grepl("shot",df$quotes)

#How many quotes have the word "shot" in them?
sum(str_detect(df$quotes,"shot"))
sum(grepl("shot",df$quotes))

#How many times does the word "shot" come up in all the texts together?
str_count(df$quotes,"shot")
sum(str_count(df$quotes,"shot"))
```






## Searching and Filtering
```{r}
#Searching for authors
df = rbind(df,data.frame(quotes = "\"Success is peace of mind which is a direct result of self-satisfaction in knowing you did your best to become the best you are capable of becoming.\" - John Wooden"))
df = rbind(df,data.frame(quotes = "\"In a gentle way, you can shake the world.\" - Mahatma Gandhi"))
df

sum(str_detect(df$quotes,"John Wooden"))
#or
sum(grepl("John Wooden",df$quotes))

#Return the quotes from Wooden

writeLines(df$quotes[str_detect(df$quotes,"John Wooden")])
#or
writeLines(df$quotes[grepl("John Wooden",df$quotes)])
#or
df %>% filter(str_detect(quotes,"John Wooden"))
```


#Search for Matches and a quick plot
```{r}
colors = c("orange","blue","yellow","green","purple","brown","red")
color_expression = str_c(colors, collapse = "|")
color_expression
has_color = str_subset(sentences,color_expression)
has_color
matches = str_extract(has_color,color_expression)
matches
matches_all = str_extract_all(has_color,color_expression, simplify = TRUE)
matches_all
class(matches_all)
matches_all = unlist(str_extract_all(has_color,color_expression))
matches_all
matchDF = data.frame(Colors = matches_all)
matchDF %>% ggplot(aes(x = Colors, fill = Colors)) + geom_bar()
matchDF %>% ggplot(aes(x = Colors, fill = Colors)) + geom_bar()+ scale_fill_manual(values=colors)
colors[order(colors)]
matchDF %>% ggplot(aes(x = Colors, fill = Colors)) + geom_bar()+ scale_fill_manual(values=colors[order(colors)])
```

## Grouped Matches
```{r}
author = "( -| - )([^ ]{2,}) ([^ ]{2,})"
authors = df$quotes

authors %>% str_extract(author)
authors %>% str_match(author)

dfAuthors = data.frame(authors %>% str_match(author)) 
names(dfAuthors) = c("Full","Dash","First","Last")
dfAuthors
dfAuthors %>% select(c(First,Last))
```

## Replacement
```{r}
roster = data.frame(Name = c("John", "Nancy", "Fred", "Sam", "Julie"), Gender = c("male", "F", "M", "Female", "female"), Major = c("?","Math","Comp Sci", "?", ""))
str_replace(roster$Gender,"(male|M)", "Male")
str_replace(roster$Gender,"(\\bmale\\b|\\bM\\b)", "Male")
roster$Gender = str_replace(roster$Gender,"(\\bmale\\b|\\bM\\b)", "Male")
roster$Gender = str_replace(roster$Gender,"(\\bfemale\\b|\\bF\\b)", "Female")





#You Try Change all the ? marks and null spaces to "<NA>".
#Don't look at the next line of code yet :)







str_replace(roster$Major,"(\\?|)", "NA")

roster$Major = str_replace(roster$Major,"(\\?)", "")
roster

str(roster)

roster$Gender = as.factor(roster$Gender)
roster$Major = as.factor(roster$Major)
levels(roster$Major) = droplevels(roster$Major,"")
str(roster)
```

##Split
```{r}
#Basic Demo of str split

str_split("Muhatma_Gandhi, John_Wooden, Maya_Angelou" , "(_|,)")
str_split("Muhatma_Gandhi, John_Wooden, Maya_Angelou" , "(_|, )")

#Look into the first object of the list (which is a string of strings)
str_split("Muhatma_Gandhi, John_Wooden, Maya_Angelou" , "(_|, )")[[1]] 
#or
unlist(str_split("Muhatma_Gandhi, John_Wooden, Maya_Angelou" , "(_|, )")) 

#Don't confuse with separate()
library(tidyr)
Authors = data.frame(Author = c("Muhatma_Gandhi", "John_Wooden", "Maya_Angelou"))
Authors %>% separate(Author,into = c("First", "Last"), sep = "_")
```

##Split string into words
```{r}
s1 = sentences[1]
str_split(s1," ")
str_split(s1,"\\b")
str_split(s1,boundary("word"))
unlist(str_split(s1,boundary("word")))

s2 = "I like chocolate. It is my favorite."
str_split(s2," ")
str_split(s2,boundary("word"))
```

#regex()
```{r}
roster = data.frame(Name = c("John", "Nancy", "Fred", "Sam", "Julie", "Pat", "Mel", "Jay"), Gender = c("male", "F", "M", "Female", "femaLe","m","f","MaLe"), Major = c("NA","Math","Comp Sci", "NA", "NA", "NA", "English", "EE"))
roster
str(roster)

roster$Gender = str_replace(roster$Gender,"(\\bmale\\b|\\bM\\b)", "Male")
roster$Gender = str_replace(roster$Gender,"(\\bfemale\\b|\\bF\\b)", "Female")
roster
str(roster)

roster$Gender = str_replace(roster$Gender,regex("(\\bmale\\b|\\bM\\b)",ignore_case = TRUE), "Male")
roster$Gender = str_replace(roster$Gender,regex("(\\bfemale\\b|\\bF\\b)",ignore_case = TRUE), "Female")
roster$Gender = as.factor(roster$Gender)
roster
str(roster)

roster[roster$Major=="NA",'Major'] <- NA
roster

```





#Example NYT

```{r NYT}
######################

# Loading the Data from the NYT API

#Information about NYT API
#https://developer.nytimes.com/faq#a11
######################

library(tidyr)
library(plyr)
library(jsonlite)
library(dplyr)
library(tidyverse)

NYTIMES_KEY = "OG89fUubcS8FXofVrLA4dmIOHh5omiFa" #Your Key Here … get from NTY API website

# Let's set some parameters
term <- "Central+Park+Jogger" # Need to use + to string together separate words
begin_date <- "19890419"
end_date <- "19910419"

baseurl <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",term,
                  "&begin_date=",begin_date,"&end_date=",end_date,
                  "&facet_filter=true&api-key=",NYTIMES_KEY, sep="")

baseurl

initialQuery <- jsonlite::fromJSON(baseurl)
maxPages <- round((initialQuery$response$meta$hits[1] / 10)-1)

maxPages

pages <- list()
for(i in 0:maxPages){
  nytSearch <- jsonlite::fromJSON(paste0(baseurl, "&page=", i), flatten = TRUE) %>% data.frame() 
  message("Retrieving page ", i)
  pages[[i+1]] <- nytSearch 
  Sys.sleep(7) # to avoid http 429 error (too many requests)
}

allNYTSearch <- rbind_pages(pages)


#Segmentation

# Visualize coverage by section
allNYTSearch %>% 
  group_by(response.docs.type_of_material) %>%
  dplyr::summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=response.docs.type_of_material, fill=response.docs.type_of_material), stat = "identity") + coord_flip()

#Make another column of News versus Other ... The labels

allNYTSearch$NewsOrOther = ifelse(allNYTSearch$response.docs.type_of_material == "News","News","Other")
#There is an NA in NewsOrOther

# Visualize coverage of News or Other
allNYTSearch[!is.na(allNYTSearch$NewsOrOther),] %>% 
  group_by(NewsOrOther) %>%
  dplyr::summarize(count=n()) %>%
  mutate(percent = (count / sum(count))*100) %>%
  ggplot() +
  geom_bar(aes(y=percent, x=NewsOrOther, fill=NewsOrOther), stat = "identity") + coord_flip()



#Identify Article and Tokenize

#Article 4 is "Other"
#Artilce 9 is "News"
#Artilce 44 is "Other"

ArticleToClassify = allNYTSearch[4,] # Check out 4, 9 and 44
ArticleToClassify$response.docs.headline.main

trueType = ArticleToClassify$NewsOrOther[1]
trueType

library(tm) #text mining library provides the stopwords() function
stopwords()

#The [^[:alnum:] ] replaces all non alphanumeric characters with nulls.  
theText = unlist(str_split(str_replace_all(ArticleToClassify$response.docs.headline.main,"[^[:alnum:] ]", ""), boundary("word"))) #Take out all but alpha numeric characters from search string

theText

wordsToTakeOut = stopwords()

#put word boundaries stopwords so that we don't detect partial words later
wordsToTakeOut = str_c(wordsToTakeOut,collapse = "\\b|\\b") 
wordsToTakeOut = str_c("\\b",wordsToTakeOut,"\\b")
wordsToTakeOut

importantWords = theText[!str_detect(theText,regex(wordsToTakeOut,ignore_case = TRUE))]

importantWords

#Find Percentages in News and Other

newsArticles = allNYTSearch %>% filter(NewsOrOther == "News")
otherArticles = allNYTSearch %>% filter(NewsOrOther == "Other")

numNewsArticles = dim(newsArticles)[1]
numOtherArticles = dim(otherArticles)[1]

numNewsArticles
numOtherArticles

thePercentHolderNews = c()
thePercentHolderOther = c()

for(i in 1 : length(importantWords)) #for each important word in the headline
{
  #number of News articles that have the ith word in the headline of interest
  numNews = sum(str_count(newsArticles$response.docs.headline.main,importantWords[i]))
  #number of Other articles that have the ith word in the headline of interest
   numOther = sum(str_count(otherArticles$response.docs.headline.main,importantWords[i]))
 
  #percentage of News articles that have the ith word in the headline of interest 
  thePercentHolderNews[i] = numNews / numNewsArticles
  #percentage of Other articles that have the ith word in the headline of interest
  thePercentHolderOther[i] = numOther / numOtherArticles
  
  #all the News percentages (for each word)
  thePercentHolderNews
  #all the Other percentages (for each word)
  thePercentHolderOther
  
}

thePercentHolderNews
thePercentHolderOther

classifiedAs = if_else(sum(thePercentHolderNews)>sum(thePercentHolderOther),"News","Other")
sum(thePercentHolderNews)
sum(thePercentHolderOther)

Result = str_c("The ", trueType," article was classified as ", classifiedAs, " with a News score of: ",round(sum(thePercentHolderNews),4), " and an Other score of: ", round(sum(thePercentHolderOther),4), ".") 
Result


## VISUALIZE

articleStats = data.frame(Word = importantWords, newsScore = thePercentHolderNews, otherScore = thePercentHolderOther)

# Wide Form / Not Tidy
articleStats

#Tidy and Plot
articleStats[,c(2,3)] %>% gather(Type,Percent) %>% mutate(Word = rep(articleStats$Word,2)) %>% ggplot(aes(y = Percent, x = Type, fill = Word)) + geom_col()

articleStats[,c(2,3)] %>% gather(Type,Percent) %>% mutate(Word = rep(articleStats$Word,2)) %>% ggplot(aes(y = Percent, x = Type, fill = Word)) + geom_col() + facet_wrap(~Word)
```







