---
title: "Unit 5 For Live Session Solution"
author: "Bivin"
date: "9/23/2019"
output: html_document
---


# Unit 5 Live Session R Code

## Part 1 BBall

Creating the TotalInches column:
```{r}
BBall = read.csv(file.choose(),header = TRUE)
BBall = BBall %>% separate(height, c("feet","inches")) %>% mutate(TotalInches = 12*as.numeric(feet) + as.numeric(inches))
head(BBall)

BBall = BBall %>% separate(height, c("feet","inches")) %>% mutate(TotalInches = 12*as.numeric(feet) + as.numeric(inches))
```

Making the histogram:
```{r}
BBall %>% ggplot(aes(x = TotalInches, color = position)) + geom_histogram() + facet_wrap(~position)
```

## Part 2 Soccer (Football)

Creating the To
```{r}
FIFA = read.csv(file.choose(),header = TRUE)

FIFA = FIFA %>% separate(Height, c("feet","inches","'")) %>% mutate(TotalInches = 12*as.numeric(feet) + as.numeric(inches)) %>% mutate(WeightNum = as.numeric(str_sub(FIFA$Weight,1,3)))
head(FIFA$WeightNum)

#Another way for weight:
FIFA = FIFA %>% separate(Height, c("feet","inches","'")) %>% mutate(TotalInches = 12*as.numeric(feet) + as.numeric(inches)) %>% mutate(WeightNum2 = as.numeric(str_replace(FIFA$Weight, "lbs", "")))
head(FIFA$WeightNum2)

#or 

FIFA = FIFA %>% separate(Height, c("feet","inches","'")) %>% mutate(TotalInches = 12*as.numeric(feet) + as.numeric(inches)) %>% mutate(WeightNum3 = as.numeric(str_replace(FIFA$Weight, "...$", "")))
head(FIFA$WeightNum3)



```

There appears to be a strong positive and linear relationship beween height and weigth (as you would expect).  But is it the same for positions LM and LB?
```{r}
FIFA %>% ggplot(aes(x = WeightNum, y = TotalInches)) + geom_point() + ggtitle("Height v. Weight")
```

Visually, there is not a lot of evidence to suggest the  relationship is different. 
```{r}
FIFA %>% filter(Position == "LB" | Position == "LM") %>% ggplot(aes(x = WeightNum, y = TotalInches, color = Position)) + geom_point() + geom_smooth(method = "lm") + ggtitle("Height v. Weight (LM vs. LB)") +facet_wrap(~Position) + xlab("Weight (lbs)") + ylab("Height (inches)")

```

## Part 3: Baby Names

```{r}
# 1A
# setwd()
df <- read.table(file.choose(), sep=";", header=FALSE, col.names = c("Name", "Gender", "BornThisYear"))

# 1B
summary(df)
str(df)

#1C
df$Name[grep("y{3}", df$Name)] # There are others, of course; this regular expression does not even say it's at the end of the word

#1D 
y2016 <- df[-grep("y{3}", df$Name),] # You could also hard-code which row it is, but I recommend this.

# 2A
y2015 <- read.table(file.choose(), sep=",", header=FALSE, col.names = c("Name", "Gender", "BornThisYear"))
# Notice that my columns are the same name.  If students have different names, the code below changes a little.


#2B
tail(y2015,n=10)
# For 'interesting,' it could be a lot of things.  For one, all 10 are male.  All 10 start with the letter Z.  All are unpopular names. All are male!  

#2C
final<-merge(y2016, y2015, by=c("Name", "Gender"), all=FALSE) # There are other combinations of this.  Inner Join

#3A
final$Total <- final$BornThisYear.x + final$BornThisYear.y # Creating column
sum(final$Total) # Figuring out everyone in it

#3B
a<-data.frame(final[order(-final$Total), c("Name", "Gender","Total")]) # You could also do tail here if you did ascending order.
head(a, n=10)

b<- final %>% arrange(desc(Total))
head(b, n=10)


#3C
# Lots of strategies here.  You could include only girls, omit only boys, use group bys, etc.  Use your judgment as grader.
girl<-head(a[a$Gender!="M",], n=10)
# It doesn't have to be as elegant a solution as this.  I just care that they get around here.

girl2 = head(b %>% filter(Gender == "F"), n = 10)


# 3D
girl<-girl[,c("Name", "Total")]
girl2 = girl2 %>% select(Name,Total)
write.csv(girl2, "itsagirl.csv", row.names = FALSE)


# 4 Visualization

girl$Name = droplevels(girl$Name)

girl %>% ggplot(aes(x = factor(Name, levels = Name[order(Total)]), y = Total, color = Name)) + geom_col() + ggtitle("Top 10 Girl Names!") 

boy<-head(a[a$Gender == "M",], n=10)
boy2 = head(b %>% filter(Gender == "M"))

boy$Name = droplevels(boy$Name)

boy %>% ggplot(aes(x = factor(Name, levels = Name[order(Total)]), y = Total, color = Name)) + geom_col() + ggtitle("Top 10 Boy Names!") 

```
